Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2016-01-08 paper/paper_acm.tex
remove LT

Nevertheless, the adaptation of the sheafification process in higher
topos theory to homotopy type theory is not straightforward because
the construction in higher toposes is restricted to the initial
Grothendieck setting which is still very topologically oriented, and
not very amenable to formalization in type theory. It seems more
promising to use a generalized notion of sheafification, called
Lawvere-Tierney sheafification~\cite{tierney1972sheaf,maclanemoerdijk}
(LT-sheafification), but this construction has not been considered yet
in the setting of higher topos theory. This rises two issues that this
paper addresses: (i) how to lift the notion of LT-sheafification to a
higher dimensional setting and (ii) is it possible to formalize this
new definition in homotopy type theory.

@@

Technically, LT-sheafification in homotopy type theory allows to
construct a model of homotopy type theory (restricted to finite homotopy
types) with new logical principles (such as EM) from a model of
homotopy type theory.

That is, LT-sheafification can be seen as a model transformation that
allows to justify new axioms in homotopy type theory. 
%
But as this transformation is internalized inside homotopy type theory
itself, it also provides a computational content to the added axiom,
making them more amenable to be used in practice inside a proof
assistant.

@@

The work on forcing in type theory~\cite{jaber2012extending} also
gives a computational meaning to a type theory enriched with new
logical or computational principle, but for other kinds of principles
as the ones interpreted by LT-sheafification.
%
Actually, forcing is entirely complementary to LT-sheafification as
forcing corresponds to the presheaf construction while
LT-sheafification corresponds to the topological transformation that
allows to go from the presheaf construction to the sheaf construction.

@@

Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the
construction.
% 
Section~\ref{sec:hott} presents the foundations of homotopy type theory
and presents a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
%
Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
modalities, on which this paper is based.
%
Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
type theory.
%
Section~\ref{sec:sheafification} explains the construction of the
LT-sheafification functor.
%
The definitions of this paper use an informal presentation of
homotopy type theory, working entirely inside of Univalent
Foundations. 
 %
Section~\ref{sec:extension-type} describes how to use this
construction to define a modality on $\Type$ from a modality on
$\HProp$.
%
Section~\ref{sec:formalization} discusses the details of the
formalization in Coq using formal type theory, in particular issues
concerning universe polymorphism.
%
Finally, Section~\ref{sec:future-works} concludes on future works.

@@

In order to understand the construction of LT-sheafification in homotopy
type theory, we first briefly recall the corresponding construction in
topos theory and then gives an overview of our construction.

@@

Furthermore, one can define a left adjoint to the inclusion, the
LT-sheafification functor
%
\begin{mymath}
a_j : \E \rightarrow \Sh{\E}
\end{mymath}%
which exhibits $\Sh{\E}$ as a reflective
subcategory (or reflective localization) of $\E$. This
means that logical principles valid in $\E$ are still valid in
$\Sh{\E}$.

@@

To extend Lawvere-Tierney sheafification to homotopy type theory, the
first thing to understand is that the construction cannot be done in one
single step anymore. It must rather be performed by induction on the
level of homotopy types. More precisely, the first layer of
LT-sheafification is defined for $\Type_{0}$, given a topology on
$\Type_{-1}$. This part corresponds to LT-sheafification for topoï. Then,
assuming that the LT-sheafification has been constructed above level $n$,
one can define the LT-sheafification for $\Type_{n+1}$.
%
This inductive process in the definition of LT-sheafification is similar
to construction in the Grothendieck setting, where the
sheafification is obtained by applying the plus-construction $n\!+\!1$
times to
$n$-types~\cite[Chapter~6]{lurie}. Unfortunately, some types does not
have a finite homotopy level, hence this construction cannot reach
every type.

This inductive step requires to formalize the notion of left-exact
reflective subuniverse, which corresponds to a stratified version of
left-exact modality introduced in~\cite[Chapter~7]{hottbook}. A
Lawvere-Tierney topology can thus be seen as a left-exact modality on
$\Type_{-1}$ and LT-sheafification as an inductive process that extends
it as a left-exact modality on $\Type_{n}$ for any $n$.
Our inductive definition of these modalities will be given by
\[ \begin{array}{l}
   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
   \\
    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\

@@

Compare to the definition of LT-sheafification in topos theory, the main
change comes from the fact that in homotopy type theory, there is not
a single object classifier but rather an $n$-object classifier for
each $n\geq -2$ which classifies functions with $n$-truncated homotopy
fibers~\cite{sets_in_hott}. This means that when monomorphisms are used in topos-theoretic definition, the correct generalization may be $n$-truncated homotopy
fibers or embeddings (\ie $(-1)$-truncated homotopy
fibers) depending on the purpose of the definition.

For instance, the notion of separated objects (which is a technical
notion used in the construction of the LT-sheafification functor), \ie
objects for which the morphism
(\ref{equ:sheaf_def}) is just an embedding, is generalized in two
different ways.  
%
The subobjects to be considered are given by $n$-truncated homotopy fibers,
while the morphism (\ref{equ:sheaf_def}) is still required to be an
embedding.

Another technical difficulty is the fact that LT-sheafification in topos
theory uses the property that epimorphisms are coequalizers of their
kernel pairs. This property does not hold anymore in higher topoï, but
using Giraud-Rezk-Lurie axioms of $\infty$-topoï, the same property
holds by replacing kernel pairs by \v{C}ech nerves in the statement.

@@

Such a construction can be thought in two ways:
\begin{itemize}
\item as a translation from one type theory into another type theory
  satisfying new principles
\item or as a transformation, taking a model of type theory, and
  producing another model, where the desired principle is true. This
  can lead to consistency results.
\end{itemize}

@@

In this section, we review some basic definitions in homotopy type
theory that are central in our formalization but not specific to
LT-sheafification. 
% %
% Section~\ref{ssec:hott} introduces the notion of homotopy types and
% classifying objects as defined in~\cite{sets_in_hott}.
% %
% Section~\ref{sec:epi-mono-fact} presents the usual epi-mono
% factorization system in the light of homotopy type theory.
% %
% Section~\ref{sec:colim-homot-type} introduces the notion of colimits
% in homotopy type theory. 
% %
% And Section~\ref{sec:giraud-ax} defines an analogue of
% Giraud-Rezk-Lurie axioms on higher topoï that we need in the
% definition of sheafification.
% %
The definitions of Section~\ref{ssec:hott} 
are part, or direct applications of~\cite{hottbook}, while other definitions
and theorems are specific to our formalization.

@@

@@

\subsection{Colimits in Homotopy Type Theory }
\label{sec:colim-homot-type}

One desired property we would like to consider concerns the colimits
of \v{C}ech nerves (Section~\ref{sec:giraud-ax}). This section presents
a definition of colimits in a type theoretic setting.
% In the definition of (a special case of) Giraud-Rezk-Lurie axiom given in
% Section~\ref{sec:giraud-ax}, the colimit of the {\em \v{C}ech nerve}
% plays a central role.
% 
Following the definition of graphs and diagrams defined
in~\cite{lumsdaine}, we recall the definition of colimits of
diagrams overs graphs presented in~\cite{sets_in_hott}. 
%
% The main difference between limits and colimits is that limits are
% simply given by $\Sigma$-types, and thus exist already in traditional
% type theory, whereas the situation is more complicated for colimits as
% it requires the use of higher inductive types.

@@

\subsection{On Giraud-Rezk-Lurie axioms}
\label{sec:giraud-ax}

@@

The notion of left exact modalities plays a central role in this paper
because Lawvere-Tierney topologies correspond exactly to left exact
modalities at level $-1$ and LT-sheafification is described as an inductive
process that produces a left exact modality at level $n+1$ from a
level exact modality at level $n$.

@@

As the definition of LT-sheafification is based on subobject classifiers
and thus on homotopy fibers, we need to take a closer look at the way left
exact modalities preserve homotopy fibers. The following is a part of
the library~\cite{hottlib}.
%
\begin{prop}
\label{sec:defin-basic-prop}
For any $n$-truncated types $X$ and $Y$,
and any map $f:X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
is the fiber of $\modal f$ above $\eta_Y y$:
\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
(\modal f(x) = \eta_Y(y)),\end{mymath}%
where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.

@@

In this paper, we use the double negation modality at level $-1$ only,
and then extend it to higher homotopy types using LT-sheafification. This
way, we can add classical reasoning in the type theory without
collapsing it to mere propositions.

@@

As explained in in Section~\ref{sec:lawv-tiern-sheaf}, a
Lawvere-Tierney topology for a topos $\E$ corresponds to a left exact
modality at level $-1$
%
and LT-sheafification corresponds to lifting this modality to a left
exact modality at level $0$.
%
In homotopy type theory, it is possible to use similar ideas to lift
any left exact modality at level $-1$ to a left exact modality at level
$n$ using LT-sheafification, by induction on $n$.

@@

The LT-sheafification in topos theory described in
Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
%
From a left exact modality on the internal logic (a Lawvere-Tierney
topology), a new left exact modality is defined on the whole topos.
%
This construction may be seen as the first induction step between
$\HProp$ and $\Type_0$.

@@

For a given induction step, the LT-sheafification is defined in two
steps ; the first one construct, for any $T:\Type_{n+1}$ a separated
type $\separated T$, and the second one construct a sheaf given any
separated type.
% %
% \begin{enumerate}[(i)]
% \item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
%   its {\em free} separated object $\separated T$ by ``removing'' what
%   prevents $T$ from being a sheaf.
% \item {\em completion:} We add what is missing for the free 
%   separated type to be a sheaf by using closure in a sheaf.
% \end{enumerate}

@@

Again, this lemma is an adaptation of the LT-sheafification process in
topos theory, where the kernel pair of $\mu_T$
is replaced by its iterated kernel pair.

@@

\begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
  As $\mu_T$ is an surjection, using the theorem~\ref{cech} introduced in Section~\ref{sec:giraud-ax},
  we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 
  % 
  Thus the result follows from the fact that $\|C(\mu_T)\|_{n+1} =
  \mathring T$,
  which we now show.

@@

As this modality is just a step in the construction, we do not need to
show that it is left exact, we will only do it for the LT-sheafification
modality.

@@

By proposition~\ref{prop:consistent}, these left-exact modalities
induces a consistent type theory. Furthermore, the univalence remains
true in this new type theory in the following sense:
\begin{prop}
  Let $n$ be a given truncation index, and $\modal$ the modality
  associated to $n+1$ as defined in lemma~\ref{lem:type}. Then, for
  any type $A,B:\Type_n^\modal$, the canonical arrow
  $$\varphi~:~A = B \to A\simeq B$$
  is an equivalence.
\end{prop}
\begin{proof}
  The first thing to notice is that, if $X$ and $Y$ are modal, and
  $f:X \to Y$, then the mere proposition $\IsEquiv f$ is also modal.
  Therefore, it suffices to show that both $A=B$ and $A\simeq B$ are
  modal. As $\Type_n^\modal$ is modal, $A=B$ is modal. 
  Moreover, $(A\simeq B) \simeq \sum_{f:A\to B} \IsEquiv
  f$. Therefore, as $A$ and $B$ are modal, $A\simeq B$ is too.

@@

We can view sheafification in terms of model of type theory ; let
$\mathfrak M$ be a model of homotopy type theory. Using sheafification
associated to a truncation index $n$, one can build a modality
$\modal$, which is left-exact with respect to types at most
$n$-truncated. Thus, there is a model $\modal \mathfrak M$ of type
theory (using results in section~\ref{sec:new-type-theory}), where
excluded middle is true, and which is univalent when restricted to $n$-types.

@@

One issue with universe polymorphism lies in the management of
recursive definitions. Indeed, the following recursive definition of
LT-sheafification
%
% \vspace{-0.05em}
\[ \begin{array}{l}
   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
   \\
    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\

      \displaystyle{\modal_{n+1}(T)} \defeq  
      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
      \left\|
      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
% \vspace{-0.3em}
\]
%
is not allowed. 
%
This is because Coq forces the universe of the first $\Type_n$
occurring in the definition to be the same for every $n$, whereas the
universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
least one level higher as the one of $\Type_n$ occurring in
$\modal_{n}$ because of the use of $\Sigma$-type over
$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
%
% \nt{explain why it is an issue.}
%
Thus, the induction step presented in this paper has been formalized,
but the complete recursive LT-sheafification can not be defined for the
moment.
%
Note that the same increasing in the universe levels occurs in the
Rezk completion for categories~\cite{rezk}. In the definition of the
completion, they use the Yoneda embedding and representable functors,
which is similar to our use of characteristic functions.

@@

In this paper, we have defined a way to extend homotopy type theory
with new principles, leveraging Lawvere-Tierney sheafification to
higher topos theory. 
%
Using LT-sheafification, it is possible to add new logical principles to
homotopy type theory, such as the law of excluded middle, and this
without axiom.
%
Beyond this result, our work is part of a more general line of work
which aim to illustrate that homotopy type theory is a very promising
theory for formalizing mathematics inside a proof assistant.

>>>>

Nevertheless, the adaptation of the sheafification process in higher
topos theory to homotopy type theory is not straightforward because
the construction in higher toposes is restricted to the initial
Grothendieck setting which is still very topologically oriented, and
not very amenable to formalization in type theory. It seems more
promising to use a generalized notion of sheafification, called
Lawvere-Tierney
sheafification~\cite{tierney1972sheaf,maclanemoerdijk},
 but this construction has not been considered yet
in the setting of higher topos theory. This rises two issues that this
paper addresses: (i) how to lift the notion of Lawvere-Tierney sheafification to a
higher dimensional setting and (ii) is it possible to formalize this
new definition in homotopy type theory.

@@

Technically, Lawvere-Tienrey sheafification in homotopy type theory allows to
construct a model of homotopy type theory (restricted to finite homotopy
types) with new logical principles (such as EM) from a model of
homotopy type theory.

That is, it can be seen as a model transformation that
allows to justify new axioms in homotopy type theory. 
%
But as this transformation is internalized inside homotopy type theory
itself, it also provides a computational content to the added axiom,
making them more amenable to be used in practice inside a proof
assistant.

@@

The work on forcing in type theory~\cite{jaber2012extending} also
gives a computational meaning to a type theory enriched with new
logical or computational principle, but for other kinds of principles
as the ones interpreted by Lawvere-Tierney sheafification.
%
Actually, forcing is entirely complementary to Lawvere-Tierney sheafification as
forcing corresponds to the presheaf construction while
Lawvere-Tierney sheafification corresponds to the topological transformation that
allows to go from the presheaf construction to the sheaf construction.

@@

Section~\ref{sec:lawv-tiern-nutshell} presents an overview of the
construction.
% 
Section~\ref{sec:hott} presents the foundations of homotopy type theory
and presents a analogous of~\cite[Corollary 6.2.3.5]{lurie}. 
%
Section~\ref{sec:lexmod} introduces the notion of left-exact stratified
modalities, on which this paper is based.
%
Section~\ref{sec:sheaves} defines Lawvere-Tierney sheaves in homotopy
type theory.
%
Section~\ref{sec:sheafification} explains the construction of the
Lawvere-Tierney sheafification functor.
%
The definitions of this paper use an informal presentation of
homotopy type theory, working entirely inside of Univalent
Foundations. 
 %
Section~\ref{sec:extension-type} describes how to use this
construction to define a modality on $\Type$ from a modality on
$\HProp$.
%
Section~\ref{sec:formalization} discusses the details of the
formalization in Coq using formal type theory, in particular issues
concerning universe polymorphism.
%
Finally, Section~\ref{sec:future-works} concludes on future works.

@@

In order to understand the construction of Lawvere-Tierney sheafification in homotopy
type theory, we first briefly recall the corresponding construction in
topos theory and then gives an overview of our construction.

From now on, {\em sheafification} is to be understood as {\em
  Lawvere-Tierney sheafification}.

@@

Furthermore, one can define a left adjoint to the inclusion, the
sheafification functor
%
\begin{mymath}
a_j : \E \rightarrow \Sh{\E}
\end{mymath}%
which exhibits $\Sh{\E}$ as a reflective
subcategory (or reflective localization) of $\E$. This
means that logical principles valid in $\E$ are still valid in
$\Sh{\E}$.

@@

To extend Lawvere-Tierney sheafification to homotopy type theory, the
first thing to understand is that the construction cannot be done in one
single step anymore. It must rather be performed by induction on the
level of homotopy types. More precisely, the first layer of
sheafification is defined for $\Type_{0}$, given a topology on
$\Type_{-1}$. This part corresponds to sheafification for topoï. Then,
assuming that the sheafification has been constructed above level $n$,
one can define the sheafification for $\Type_{n+1}$.
%
This inductive process in the definition of sheafification is similar
to construction in the Grothendieck setting, where the
sheafification is obtained by applying the plus-construction $n\!+\!1$
times to
$n$-types~\cite[Chapter~6]{lurie}. Unfortunately, some types does not
have a finite homotopy level, hence this construction cannot reach
every type.

This inductive step requires to formalize the notion of left-exact
reflective subuniverse, which corresponds to a stratified version of
left-exact modality introduced in~\cite[Chapter~7]{hottbook}. A
Lawvere-Tierney topology can thus be seen as a left-exact modality on
$\Type_{-1}$ and sheafification as an inductive process that extends
it as a left-exact modality on $\Type_{n}$ for any $n$.
Our inductive definition of these modalities will be given by
\[ \begin{array}{l}
   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
   \\
    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\

@@

Compare to the definition of sheafification in topos theory, the main
change comes from the fact that in homotopy type theory, there is not
a single object classifier but rather an $n$-object classifier for
each $n\geq -2$ which classifies functions with $n$-truncated homotopy
fibers~\cite{sets_in_hott}. This means that when monomorphisms are used in topos-theoretic definition, the correct generalization may be $n$-truncated homotopy
fibers or embeddings (\ie $(-1)$-truncated homotopy
fibers) depending on the purpose of the definition.

For instance, the notion of separated objects (which is a technical
notion used in the construction of the sheafification functor), \ie
objects for which the morphism
(\ref{equ:sheaf_def}) is just an embedding, is generalized in two
different ways.  
%
The subobjects to be considered are given by $n$-truncated homotopy fibers,
while the morphism (\ref{equ:sheaf_def}) is still required to be an
embedding.

Another technical difficulty is the fact that sheafification in topos
theory uses the property that epimorphisms are coequalizers of their
kernel pairs. This property does not hold anymore in higher topoï, but
using Giraud-Rezk-Lurie axioms of $\infty$-topoï, the same property
holds by replacing kernel pairs by \v{C}ech nerves in the statement.

@@

Such a construction can be thought in two ways:
\begin{itemize}
\item as a translation from one type theory into another type theory
  satisfying new principles
\item or as a transformation, taking a model of type theory, and
  producing another model, where the desired principle is true. As in
  set theory or topos theory, it could lead to independence results.
\end{itemize}

@@

In this section, we review some basic definitions in homotopy type
theory that are central in our formalization but not specific to
sheafification. 
% %
% Section~\ref{ssec:hott} introduces the notion of homotopy types and
% classifying objects as defined in~\cite{sets_in_hott}.
% %
% Section~\ref{sec:epi-mono-fact} presents the usual epi-mono
% factorization system in the light of homotopy type theory.
% %
% Section~\ref{sec:colim-homot-type} introduces the notion of colimits
% in homotopy type theory. 
% %
% And Section~\ref{sec:giraud-ax} defines an analogue of
% Giraud-Rezk-Lurie axioms on higher topoï that we need in the
% definition of sheafification.
% %
The definitions of Section~\ref{ssec:hott} 
are part, or direct applications of~\cite{hottbook}, while other definitions
and theorems are specific to our formalization.

@@

Section~\ref{ssec:hott} will present homotopy levels and object
classifiers, section~\ref{ssec:colim-homot-type} introduces a theory
of colimits in homotopy type theory, illustrated by an important
example in section~\ref{ssec:giraud-ax}.

@@

\subsection{Colimits in Homotopy Type Theory }
\label{ssec:colim-homot-type}

One desired property we would like to consider concerns the colimits
of \v{C}ech nerves (Section~\ref{ssec:giraud-ax}). This section presents
a definition of colimits in a type theoretic setting.
% In the definition of (a special case of) Giraud-Rezk-Lurie axiom given in
% Section~\ref{sec:giraud-ax}, the colimit of the {\em \v{C}ech nerve}
% plays a central role.
% 
Following the definition of graphs and diagrams defined
in~\cite{lumsdaine}, we recall the definition of colimits of
diagrams overs graphs presented in~\cite{sets_in_hott}. 
%
% The main difference between limits and colimits is that limits are
% simply given by $\Sigma$-types, and thus exist already in traditional
% type theory, whereas the situation is more complicated for colimits as
% it requires the use of higher inductive types.

@@

\subsection{On Giraud-Rezk-Lurie axioms}
\label{ssec:giraud-ax}

@@

The notion of left exact modalities plays a central role in this paper
because Lawvere-Tierney topologies correspond exactly to left exact
modalities at level $-1$ and sheafification is described as an inductive
process that produces a left exact modality at level $n+1$ from a
level exact modality at level $n$.

@@

As the definition of sheafification is based on subobject classifiers
and thus on homotopy fibers, we need to take a closer look at the way left
exact modalities preserve homotopy fibers. The following is a part of
the library~\cite{hottlib}.
%
\begin{prop}
\label{sec:defin-basic-prop}
For any $n$-truncated types $X$ and $Y$,
and any map $f:X \to Y$, the modalisation of fiber of $f$ above any element $y:Y$
is the fiber of $\modal f$ above $\eta_Y y$:
\begin{mymath}[-0.3em]\modal \left(\sum_{x:X}  (f(x) = y)\right) = \sum_{x:\modal X}
(\modal f(x) = \eta_Y(y)),\end{mymath}%
where $\modal f:\modal X \to \modal Y$ is defined via the equivalence {\it (iv)} and $\eta_Y$.

@@

In this paper, we use the double negation modality at level $-1$ only,
and then extend it to higher homotopy types using sheafification. This
way, we can add classical reasoning in the type theory without
collapsing it to mere propositions.

@@

As explained in in Section~\ref{sec:lawv-tiern-sheaf}, a
Lawvere-Tierney topology for a topos $\E$ corresponds to a left exact
modality at level $-1$
%
and sheafification corresponds to lifting this modality to a left
exact modality at level $0$.
%
In homotopy type theory, it is possible to use similar ideas to lift
any left exact modality at level $-1$ to a left exact modality at level
$n$ using sheafification, by induction on $n$.

@@

The sheafification in topos theory described in
Section~\ref{sec:lawv-tiern-sheaf} works in the same way. 
%
From a left exact modality on the internal logic (a Lawvere-Tierney
topology), a new left exact modality is defined on the whole topos.
%
This construction may be seen as the first induction step between
$\HProp$ and $\Type_0$.

@@

For a given induction step, the sheafification is defined in two
steps ; the first one (section~\ref{ssec:from-type-separated}) construct, for any $T:\Type_{n+1}$ a separated
type $\separated T$, and the second one
(section~\ref{ssec:separated-to-sheaf}) construct a sheaf given any
separated type.
% %
% \begin{enumerate}[(i)]
% \item {\em separation:} From any $T$ in $\Type_{n+1}$, we construct
%   its {\em free} separated object $\separated T$ by ``removing'' what
%   prevents $T$ from being a sheaf.
% \item {\em completion:} We add what is missing for the free 
%   separated type to be a sheaf by using closure in a sheaf.
% \end{enumerate}

@@

Again, this lemma is an adaptation of the sheafification process in
topos theory, where the kernel pair of $\mu_T$
is replaced by its iterated kernel pair.

@@

\begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
  As $\mu_T$ is an surjection, using the theorem~\ref{cech} introduced in Section~\ref{ssec:giraud-ax},
  we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 
  % 
  Thus the result follows from the fact that $\|C(\mu_T)\|_{n+1} =
  \mathring T$,
  which we now show.

@@

As this modality is just a step in the construction, we do not need to
show that it is left exact, we will only do it for the sheafification
modality.

@@

By proposition~\ref{prop:consistent}, these left-exact modalities
induces a consistent type theory. Furthermore, the univalence remains
true in this new type theory in the following sense:
\begin{prop}\label{prop:univalence}
  Let $n$ be a given truncation index, and $\modal$ the modality
  associated to $n+1$ as defined in lemma~\ref{lem:type}. Then, for
  any type $A,B:\Type_n^\modal$, the canonical arrow
  $$\varphi~:~A = B \to A\simeq B$$
  is an equivalence.
\end{prop}
\begin{proof}
  The first thing to notice is that, if $X$ and $Y$ are modal, and
  $f:X \to Y$, then the mere proposition $\IsEquiv f$ is also modal.
  Therefore, it suffices to show that both $A=B$ and $A\simeq B$ are
  modal. As $\Type_n^\modal$ is modal, $A=B$ is modal. 
  Moreover, $(A\simeq B) \simeq \sum_{f:A\to B} \IsEquiv
  f$. Therefore, as $A$ and $B$ are modal, $A\simeq B$ is too.

@@

We can view sheafification in terms of model of type theory ; let
$\mathfrak M$ be a model of homotopy type theory. Using sheafification
associated to a truncation index $n$, one can build a modality
$\modal$, which is left-exact with respect to types at most
$n$-truncated. Thus, there is a model $\modal \mathfrak M$ of type
theory (using results in section~\ref{sec:new-type-theory}), where
excluded middle is true, and which is univalent when restricted to
$n$-types (in the sense of proposition~\ref{prop:univalence}).

@@

One issue with universe polymorphism lies in the management of
recursive definitions. Indeed, the following recursive definition of
sheafification
%
% \vspace{-0.05em}
\[ \begin{array}{l}
   \modal : \forall \ (n : nat), \ \Type_n \to \Type_n 
   \\
    \modal_{-1\phantom{n}}(T) \defeq\lnot\lnot T \\

      \displaystyle{\modal_{n+1}(T)} \defeq  
      \displaystyle{\sum_{u:T \to \Type_n^\modal} \!\!\!\!\modal_{-1} 
      \left\|
      \sum_{a:T} u= (\lambda t,~\modal_n (a=t))
      \right\|}
    \end{array}
% \vspace{-0.3em}
\]
%
is not allowed. 
%
This is because Coq forces the universe of the first $\Type_n$
occurring in the definition to be the same for every $n$, whereas the
universe of the first $\Type_{n+1}$ occurring in $\modal_{n+1}$ should be at
least one level higher as the one of $\Type_n$ occurring in
$\modal_{n}$ because of the use of $\Sigma$-type over
$T \to \Type_n^\modal$ and equality on the return type of $\modal_n$.
%
% \nt{explain why it is an issue.}
%
Thus, the induction step presented in this paper has been formalized,
but the complete recursive sheafification can not be defined for the
moment.
%
Note that the same increasing in the universe levels occurs in the
Rezk completion for categories~\cite{rezk}. In the definition of the
completion, they use the Yoneda embedding and representable functors,
which is similar to our use of characteristic functions.

@@

In this paper, we have defined a way to extend homotopy type theory
with new principles, leveraging Lawvere-Tierney sheafification to
higher topos theory. 
%
Using sheafification, it is possible to add new logical principles to
homotopy type theory, such as the law of excluded middle, and this
without axiom.
%
Beyond this result, our work is part of a more general line of work
which aim to illustrate that homotopy type theory is a very promising
theory for formalizing mathematics inside a proof assistant.
